Rhizaria in the oligotrophic ocean exhibit clear temporal and vertical variability.

### 2024-05-02 manuscript/main.tex
update and rerender

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

% Authors	
% This hairy bit of code is just to get "and" between the last 2
% authors. See below if you don't need that
 {\large{Alex Barth*}}{\textsuperscript{1}}%
%
, 
 {\large{Leocadio Blanco-Berical}}{\textsuperscript{2}}%
%
, 
 {\large{Rod Johnson}}{\textsuperscript{2}}%
%
%
{ and \large{Joshua Stone}}%
{\textsuperscript{1}}%
%

@@

\hangindent=1em
\hangafter=1
%
{1}.~{University of South Carolina}%
%
, %
{Biological Sciences}%
%
%
, %
{700 Sumter St 401, Columbia, SC 29208}%
%
\par\hangindent=1em\hangafter=1%
%
{2}.~{Arizone State University}%
%
, %
{School of Ocean Futures}%
%
%
, %
{Bermuda Institute of Ocean Sciences 17 Biological Station, St Georges,
GE 01}%
%

@@

%%%%%%%%% Abstract
\begin{abstract}
Recently studies have shown that Rhizaria, a super-group of marine
protists have a large role in pelagic ecosystems. They are unique in
that they construct mineral tests out of silica, calcium carbonate, or
strontium sulfate. As a consequence, Rhizaria can have large impacts on
the ocean's cycling of carbon and other elements. However, less is known
about Rhizarian ecology or their role in the pelagic food-web. Some
taxa, like certain Radiolarians are mixotrophic, hosting algal
symbionts. While other taxa are flux-feeders or even predatory
carnivores. Some prior research has suggested that Rhizaria will
partition vertically in the water column, likely due to different
trophic strategies. However, very few studies have investigated their
populations over extended periods of time. In this study, we present
data investigating Rhizarian abundance and vertical distribution from
over a year of monthly cruises into the Sargasso Sea. This study
represents the first quantification of Rhizaria throughout the
mesopelagic zone in an oligotrophic system for an extended period of
time. We use this data to investigate the hypothesis that Rhizaria
taxonomic groups will partition due to trophic mode. We also investigate
how their abundance varies in accordance with a variety of environmental
parameters. Rhizaria abundance was quantified using an Underwater Vision
Profiler (UVP5), an in-situ imaging device. Ultimately, we show that
different Rhizaria taxa will have unique vertical distribution patterns.
Models relating their abundance to environmental parameters have mixed
results, yet particle concentration is a common predictive variable,
supporting the importance of heterotrophy amongst many taxa.    
\end{abstract}

@@

\end{minipage}  \end{titlepage}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

@@

Rhizaria are an extremely diverse super-group of single-celled organisms
consisting of several phyla including Retaria (foraminifera and
radiolaria) and Cercozoa. These organisms exist in a wide range of
habitats and are widely represented in plankton communities throughout
the global ocean. While the taxonomy of these organisms has recently
undergone several reclassifications (Biard, 2022a), their presence in
ocean ecosystems has been long known to oceanographers. Some of the
earliest records of their existence are from oceanographic expeditions
in the 19th century (Haekel, 1887). Rhizaria are unique members of the
plankton and protist community because they can reach large sizes (up to
several mm in diameter) and they construct intricate mineral skeletons
out of either silica, strontium, or calcium carbonate (Biard, 2022a;
Kimoto, 2015; Nakamura and Suzuki, 2015; Suzuki and Not, 2015). Despite
their noticeable morphology and global distribution, Rhizaria were
largely understudied throughout the 20th century. The bulk of modern
plankton research has focused on hard-bodied crustacea which are
numerically dominant and easily sampled with nets and preservatives.
Fragile organisms like Rhizaria, were difficult to adequately study as
they can be destroyed through standard zooplankton sampling techniques
and preserve poorly. A number of studies in the late 1900s did employ
alternative techniques to quantify Rhizaria including diaphragm pumps
(Michaels, 1988) or blue-water SCUBA collections (Bijma et al., 1990;
Caron et al., 1995; Caron and Be, 1984). However, the bulk of Rhizaria
research was constrained to sediment traps or paleontological studies of
sediment (Boltovskoy et al., 1993; Takahashi et al., 1983). Only
recently has the advent of molecular techniques and in-situ imaging
tools ignited a renewed focus on Rhizaria in pelagic ecosystems (Caron,
2016).

The wave of new data on Rhizaria has facilitated an improved
understanding of the significance in ocean ecosystem functions. Firstly,
taxonomists have been able to greatly refine the understanding of
evolutionary relationships amongst these diverse protists (Aurahs et
al., 2009; Biard et al., 2015; Cavalier-Smith et al., 2018; Decelle et
al., 2013, 2012; rev by Biard, 2022a). DNA metabarcoding studies have
revealed insights into the distributional patterns Llopis Monferrer et
al. (2022), ecological relationships (Decelle et al., 2012; Nakamura et
al., 2023), and contribution to biogeochemical fluxes (Guidi et al.,
2016; Gutierrez-Rodriguez et al., 2019). Transcriptomic and proteomic
approaches also have been used to quantify rhizarian contribution to
community metabolism (Cohen et al., 2023). Yet, despite the excellent
taxonomic resolution provided by molecular approaches, they do not
provide a truly quantitative metric for estimating Rhizarian abundance
or biomass. In-situ imaging tools however, offer the ability to observe
organisms in the natural state and quantify their abundance (Barth and
Stone, in review; Ohman, 2019). While there were early applicaitons of
imaging tools to document Rhizaria (Dennett et al., 2002), Biard et al.
(2016)'s report from a global imaging dataset highlighted the importance
of Rhizaria to the total standing stock of marine carbon. Due to their
large sizes, ability to concentrate smaller particles and the unique
structure of their mineral skeletons, Rhizaria have the potential to
massively influence ocean biogeochemical cycling. A number of studies
have made large advances in estimating the contribution of Rhizaria to
ocean cycling of carbon (Gutierrez-Rodriguez et al., 2019; Ikenoue et
al., 2019; Lampitt et al., 2009; Stukel et al., 2018), silica (Biard et
al., 2018; Llopis Monferrer et al., 2021), and strontium (Decelle et
al., 2013). Still, Rhizarian ecological roles are not well understood
(Biard, 2022a). This is a major challenge as it is critical to
understand the ecological role of plankton to fully incorporate them
into biological oceanographic models.

The ecological role of Rhizaria in plankton communities is complicated
due to the fact different taxa can exhibit every different trophic
modes. As zooplankton, rhizaria are predominately heterotrophic (Biard,
2022a), yet their feeding modes can be quite varied. Phaeodarians
(family Cercozoa) are largely thought to be flux-feeders, collecting and
feeding on sinking particles (Nakamura and Suzuki, 2015; Stukel et al.,
2019). Alternatively, Retaria can be either exclusively heterotrophic or
mixotrophic, utilizing photosynthetic algal symbionts (Anderson, 2014;
Decelle et al., 2015). Mixotrophic foraminifera host a variety of
endosymbiont partners (Decelle et al., 2015; Lee, 2006), which are
thought to support early and adult life stages and contribute to total
primary productivity (Kimoto, 2015). Still, foraminifera are omnivorous,
possibly even predominately carnivorous with several studies suggesting
that they can be effective predators (Anderson and Bé, 1976; Gaskell et
al., 2019), mainly consuming live copepods (Caron and Be, 1984).
Radiolaria have several lineages all which have some taxa who are well
known to host symbionts (Biard, 2022b). Amongst radiolarians, arguably
the most widespread are Collodaria who can be either large solitary
cells or form massive colonies, up to several meters in length (Swanberg
and Anderson, 1981). All known Collodaria species host dinoflagellate
symbionts (Biard, 2022b) and can contribute substantially to primary
productivity, particularly in oligotrophic ocean regions (Caron et al.,
1995; Dennett et al., 2002). This Collodaria-symbiont association has
been suggested as a reason for their high abundances throughout the
photic zone of oligotrophic environments globally (Biard et al., 2017,
2016). A few Acantharean (Radiolarian order) clades host algal symbionts
(Biard, 2022b; Decelle et al., 2012), notably with two clades forming an
exclusive relationship with Phaeocystis. However, globally Acantharea
are less abundant than Collodaria (Biard, 2022a) and contribute less to
total primary productivity (Michaels et al., 1995). This may be due to
the fact several clades of Acantharea are cyst-forming and strictly
heterotrophs (Biard, 2022b; Decelle et al., 2013). Furthermore, Mars
Brisbin et al. (2020) documented apparent predation behavior in
Acantharea near the surface, suggesting that there may be a large
reliance on carnivory.

Given the high abundances, yet diverse trophic strategies found among
Rhizarian taxa, it is reasonable to expect some form of niche
partitioning. A number of studies do suggest evidence for vertical
zonation between Rhizaria groups according to various trophic
strategies. Taxa-specific studies of radiolarians suggest they may be
restricted to the euphotic zone (Boltovskoy, 2017; Michaels, 1988).
Although some studies report Acantharea in deeper waters (Decelle et
al., 2013; Gutiérrez-Rodríguez et al., 2022). Phaeodarians
alternatively, are generally found in the mesopelagic where
photosynthesis cannot occur but they can feed on sinking particles
(Stukel et al., 2018). In an imaging-based study of the whole Rhizaria
community, Biard and Ohman (2020) noted clear patterns in vertical
zonation which largely corresponded to different trophic roles. In the
oligotrophic ocean, Blanco-Bercial et al. (2022) also noted that the
protist community, including Rhizaria, partition along an autotroph and
mixotroph to heterotroph gradient with increasing depth in the water
column. Yet, few studies have made direct attempts to relate rhizaria
abundances to environmental factors (Biard and Ohman, 2020). In part,
this is due to the fact few studies have been able to sample Rhizaria in
the same location over a consistent timeframe (Boltovskoy et al., 1993;
Gutiérrez-Rodríguez et al., 2022; Hull et al., 2011; Michaels et al.,
1995; Michaels, 1988). Furthermore, no studies have utilized imaging,
arguably the best method for quantifying rhizaria, consistently
throughout the full mesopelagic. Given this lack of information, there
are many unknowns with respect to Rhizarian ecology, seasonality and
phenology across different groups.

In this study, we present a comprehensive assessment of large Rhizaria
measured for multiple months (greater than 1 year) using an in-situ
imaging approach. With this dataset, we address two critical aims. 1)
Quantification of large Rhizaria throughout the epipelagic (0-200m) and
mesopelagic (200-1000m) over the course of an annual cycle. These data
were collected in the Sargasso Sea, and represents the first study of
its kind in an oligotrophic system. 2) We aim to test the hypothesis
that Rhizaria exhibit niche partitioning according to trophic roles.
This hypothesis makes several predictions, including vertical zonation,
as seen in prior studies, but also that environmental variables related
to trophic strategy will explain abundance patterns. Specifically,
autotrophic/mixotrophic taxa will correspond to variables related to
autotrophy (chl-a concentration, primary productivity, high \(O_2\) )
and other rhizaria will correspond to factors which promote heterotrophy
(particle concentration, flux, and low \(O_2\)).

@@

Data were collected in collaboration with the Bermuda Atlantic
Time-Series Study (Lomas et al., 2013; Michaels and Knap, 1996) on board
the R/V Atlantic Explorer. Cruises were conducted at approximately
monthly intervals. Rhizaria individuals were sampled using the
Underwater Vision Profiler 5 {[}UVP5; Picheral et al. (2010){]}, a tool
which is well established to accurately quantify large Rhizaria (Barth
and Stone, 2022; Biard et al., 2016; Biard and Ohman, 2020; Drago et
al., 2022; Llopis Monferrer et al., 2022; Panaïotis et al., 2023; Stukel
et al., 2019; Stukel et al., 2018). The UVP5 was mounted to the sampling
rosette and collected data autonomously on routine casts, from which
only the downcast data are utilized. The UVP5 was deployed from
June-September 2019 then from October 2020 - January 2022. Casts were
filtered to only include data collected in the BATS region, far offshore
of Bermuda into the Sargasso Sea (approximately 31.0\(^oN\)-32.5\(^oN\),
64.25\(^oW\)-63\(^oW\); Supplemental Figure 1). In general, casts
extended to either 200m, 500m, or 1200m deep, with a few extended into
the bathypelagic (4500m). However, Rhizaria were only typically found in
large abundances throughout the epipelagic and mesopelagic zones. As
such, we limit this study to results from the upper 1000m of the water
column.

A variety of biotic and abiotic data were collected during each BATS
cruise. Briefly, we will explain the data utilized in this study. The
UVP5 provided particle count data at a high-frequency from each cast.
Particle concentration was calculated from this data for all particles
184\(\mu m\) - 450\(\mu m\). The lower size range was set by what could
be reliably sampled by the UVP5's pixel resolution (\textgreater2px) and
the upper size range is representative of a potential prey field for
mesozooplankton (Whitmore and Ohman, 2021). Salinity, temperature,
Dissolved Oxygen (DO), and in-situ chlorophyll fluorescence were
measured at high-frequencies on each cast, attached to the same CTD
rosette as the UVP. On certain casts, niskin bottles were used to
collect bacterial abundance estimates (via epifluorescence microscopy)
as well as measure inorganic nutrients (\(NO_3\), and \(Si\) as
silicate/sicilic acid) at discrete depths. On each cruise, flux
estimates of total mass, carbon, and nitrogen were also collected using
sediment traps (in the present study we utilize flux to the mesopelagic
as the flux at 200m). Also primary productivity was estimated through
measuring \(C^{14}\) uptake rates from in-situ incubations. Full
descriptions of the BATS sampling program, methods, and data are
available freely online
\href{https://bats.bios.asu.edu/bats-data/}{(https://bats.bios.asu.edu/bats-data/)}
and reviewed in (Lomas et al., 2013).

Environmental data were processed in a variety of ways to match the
format of the Rhizaria abundance estimates (See below). CTD data were
collected at higher frequency than the UVP, so these data were averaged
within matching bins to the UVP5 data. Data from niskin bottles were
first linearly interpolated with a 1m resolution for each cast where
data were available. Then, because data were collected at different
frequencies over the course of a cruise, the interpolated data were
averaged across casts, then averaged into matching UVP5-sized bins.
Primary productivity estimates were totaled within the euphotic zone to
represent a ``total euphotic productivity''. Finally the data from the
sediment trap deployed to 200m was used to represent ``flux to the
mesopelagic''.

@@

Individual vignettes of Rhizaria images were identified using the
classification platform Ecotaxa (Picheral et al., n.d.). Data were
pre-sorted utilizing a random-forest classifier and pre-trained learning
set. Taxonomic classification were done based on morphology exclusively.
While there are sparse taxonomic guides for in-situ images of rhizaria,
identification largely relied on descriptions in (Nakamura and Suzuki,
2015; Suzuki and Not, 2015; and Biard and Ohman, 2020). Using the
aforementioned sources and publicly available ecotaxa projects, I
constructed a guide which can be found here:
\protect\hyperlink{0}{https://thealexbarth.github.io/media//Project\_Items/Oligotrophic\_Community/ecotaxa\_UVP-guide-stone-lab.pdf}.
Broadly, Rhizaria were classified as Foraminifera, Radiolarians
(Acantharea or Collodaria), or a variety of Phaeodarian families (Figure
1). When identification could not be confidently made between a few
candidate taxa, a less specific label was used. As a result, we have
data from ``unidentified Rhizaria'', which typically were vignettes not
distinguishable between Aulacanthidae or Acantharea or ``unidentified
Phaeodaria'', which are clearly phaeodaria but not distinguishable into
a family.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=1in,textwidth=6.5in,right=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

% Authors	
% This hairy bit of code is just to get "and" between the last 2
% authors. See below if you don't need that
 {\large{Alex
Barth*}}{\textsuperscript{1}}\textsuperscript{,}{\textsuperscript{2}}%
%
, 
 {\large{Leocadio Blanco-Berical}}{\textsuperscript{3}}%
%
, 
 {\large{Rod Johnson}}{\textsuperscript{3}}%
%
%
{ and \large{Joshua Stone}}%
{\textsuperscript{1}}%
%

@@

\hangindent=1em
\hangafter=1
%
{1}.~{University of South Carolina}%
%
, %
{Biological Sciences}%
%
%
, %
{700 Sumter St 401, Columbia, SC 29208}%
%
\par\hangindent=1em\hangafter=1%
%
{2}.~{University of Texas at Austin}%
%
, %
{Marine Science Institute}%
%
%
, %
{750 Channel Drive, Port Aransas, Texas, USA}%
%
\par\hangindent=1em\hangafter=1%
%
{3}.~{Arizone State University}%
%
, %
{School of Ocean Futures}%
%
%
, %
{Bermuda Institute of Ocean Sciences 17 Biological Station, St Georges,
GE 01}%
%

@@

%%%%%%%%% Abstract
\begin{abstract}
Recently studies have shown that Rhizaria, a super-group of marine
protists, have a large role in pelagic ecosystems. They are unique in
that they construct mineral tests out of silica, calcium carbonate, or
strontium sulfate. As a consequence, Rhizaria can have large impacts on
the ocean's cycling of carbon and other elements. However, less is known
about Rhizaria ecology or their role in the pelagic food-web. Some taxa,
like certain Radiolarians, are mixotrophic, hosting algal symbionts.
While other taxa are flux-feeders or even predatory carnivores. Some
prior research has suggested that Rhizaria will partition vertically in
the water column, likely due to different trophic strategies. However,
very few studies have investigated their populations over extended
periods of time. In this study, we present data investigating Rhizaria
abundance and vertical distribution from over a year of monthly cruises
in the Sargasso Sea. This study represents the first quantification of
Rhizaria throughout the mesopelagic zone in an oligotrophic system for
an extended period of time. We use this data to investigate the
hypothesis that Rhizaria taxonomic groups will partition due to trophic
mode. We also investigate how their abundance varies in accordance with
environmental parameters. Rhizaria abundance was quantified using an
Underwater Vision Profiler (UVP5), an in-situ imaging device.
Ultimately, we show that different Rhizaria taxa will have unique
vertical distribution patterns. Models relating their abundance to
environmental parameters have mixed results, yet particle concentration
is a common predictive variable, supporting the importance of
heterotrophy amongst many taxa.    
\end{abstract}

@@

\end{minipage}  \end{titlepage}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable]}{\end{tcolorbox}}\fi

@@

Rhizaria are an extremely diverse super-group of single-celled organisms
consisting of several phyla including Retaria (foraminifera and
radiolaria) and Cercozoa. These organisms exist in a wide range of
habitats and are widely represented in plankton communities throughout
the global ocean. While the taxonomy of these organisms has recently
undergone several reclassifications (Biard, 2022a), their presence in
ocean ecosystems has been long known to oceanographers. Some of the
earliest records of their existence are from oceanographic expeditions
in the 19th century (Haekel, 1887). Rhizaria are unique members of the
plankton and protist community because they can reach large sizes (up to
several mm in diameter) and they construct intricate mineral skeletons
out of either silica, strontium, or calcium carbonate (Biard, 2022a;
Kimoto, 2015; Nakamura and Suzuki, 2015; Suzuki and Not, 2015). Despite
their noticeable morphology and global distribution, Rhizaria were
largely understudied throughout the 20th century. The bulk of modern
plankton research has focused on hard-bodied crustacea which are
numerically dominant and easily sampled with nets and preservatives.
Fragile organisms like Rhizaria were difficult to adequately study as
they can be destroyed through standard zooplankton sampling techniques
and preserve poorly. A number of studies in the late 1900s did employ
alternative techniques to quantify Rhizaria including diaphragm pumps
(Michaels, 1988) or blue-water SCUBA collections (Bijma et al., 1990;
Caron et al., 1995; Caron and Be, 1984). However, the bulk of Rhizaria
research was constrained to sediment traps or paleontological studies of
sediment (Boltovskoy et al., 1993; Takahashi et al., 1983). Only
recently has the advent of molecular techniques and in-situ imaging
tools ignited a renewed focus on Rhizaria in pelagic ecosystems (Caron,
2016).

The wave of new data on Rhizaria has facilitated an improved
understanding of the significance in ocean ecosystem functions. Firstly,
taxonomists have been able to greatly refine the understanding of
evolutionary relationships amongst these diverse protists (Aurahs et
al., 2009; Biard et al., 2015; Cavalier-Smith et al., 2018; Decelle et
al., 2013, 2012; rev by Biard, 2022a). DNA metabarcoding studies have
revealed insights into the distributional patterns (Biard et al., 2017;
Blanco-Bercial et al., 2022; Decelle et al., 2013; Llopis Monferrer et
al., 2022; Mars Brisbin et al., 2020; Sogawa et al., 2022), ecological
relationships (Decelle et al., 2012; Nakamura et al., 2023), and
contribution to biogeochemical fluxes (Guidi et al., 2016;
Gutierrez-Rodriguez et al., 2019). Transcriptomic and proteomic
approaches also have been used to quantify Rhizaria contribution to
community metabolism (Cohen et al., 2023). Yet, despite the excellent
taxonomic resolution provided by molecular approaches, they do not
provide a truly quantitative metric for estimating Rhizaria abundance or
biomass. In-situ imaging tools however, offer the ability to observe
organisms in the natural state and quantify their abundance (Barth and
Stone, In press; Ohman, 2019). Biard et al. (2016) utilized in-situ
imaging at a global scale to suggest Rhizaria were substantial
contributors to the ocean carbon standing stock. While more recent
calculations suggest lower carbon contribution (Laget et al., 2024),
Rhizaria nonetheless have substantial influences on biogeochemical
cycling. Due to their large sizes, ability to concentrate smaller
particles and the unique structure of their mineral skeletons, Rhizaria
have the potential to massively influence ocean biogeochemical cycling.
A number of studies have made large advances in estimating the
contribution of Rhizaria to ocean cycling of carbon (Gutierrez-Rodriguez
et al., 2019; Ikenoue et al., 2019; Lampitt et al., 2009; Stukel et al.,
2018), silica (Biard et al., 2018; Llopis Monferrer et al., 2021), and
strontium (Decelle et al., 2013). Still, Rhizaria ecological roles are
not well understood (Biard, 2022a). This is a major challenge as it is
critical to understand the ecological role of plankton to fully
incorporate them into biological oceanographic models.

The ecological role of Rhizaria in plankton communities is complicated
due to the fact different taxa can exhibit every different trophic
modes. As zooplankton, rhizaria are predominately heterotrophic (Biard,
2022a), yet their feeding modes can be quite varied. Phaeodarias (family
Cercozoa) are largely thought to be flux-feeders, collecting and feeding
on sinking particles (Nakamura and Suzuki, 2015; Stukel et al., 2019).
Alternatively, Retaria can be either exclusively heterotrophic or
mixotrophic, utilizing photosynthetic algal symbionts (Anderson, 2014;
Decelle et al., 2015). Mixotrophic foraminifera host a variety of
endosymbiont partners (Decelle et al., 2015; Lee, 2006), which are
thought to support early and adult life stages and significantly
contribute to total primary productivity (Kimoto, 2015). Still,
foraminifera are omnivorous, possibly even predominately carnivorous,
with several studies suggesting that they can be effective predators
(Anderson and Bé, 1976; Gaskell et al., 2019), mainly consuming live
copepods (Caron and Be, 1984). Radiolaria have several lineages, all of
which have some taxa well known to host symbionts (Biard, 2022b).
Amongst Radiolarias, arguably the most widespread are Collodaria who can
be either large solitary cells or form massive colonies, up to several
meters in length (Swanberg and Anderson, 1981). All known Collodaria
species host dinoflagellate symbionts (Biard, 2022b) and can contribute
substantially to primary productivity, particularly in oligotrophic
ocean regions (Caron et al., 1995; Dennett et al., 2002). This
Collodaria-symbiont association has been suggested as a reason for their
high abundances throughout the photic zone of oligotrophic environments
globally (Biard et al., 2017, 2016). A few Acantharea (Radiolaria order)
clades host algal symbionts (Biard, 2022b; Decelle et al., 2012),
notably with two clades forming an exclusive relationship with
\emph{Phaeocystis}. However, globally, Acantharea are less abundant than
Collodaria (Biard, 2022a) and contribute less to total primary
productivity (Michaels et al., 1995). This may be due to the fact
several clades of Acantharea are cyst-forming and strictly heterotrophs
(Biard, 2022b; Decelle et al., 2013). Furthermore, Mars Brisbin et al.
(2020) documented apparent predation behavior in Acantharea near the
surface, suggesting that there may be a larger reliance on carnivory.

Given the high abundances, yet diverse trophic strategies found among
Rhizaria taxa, it is reasonable to expect some form of niche
partitioning. A number of studies do suggest evidence for vertical
zonation between Rhizaria groups according to various trophic
strategies. Taxa-specific studies of Radiolarias suggest they may be
restricted to the euphotic zone (Boltovskoy, 2017; Michaels, 1988).
Although some studies report Acantharea in deeper waters (Decelle et
al., 2013; Gutiérrez-Rodríguez et al., 2022), Phaeodarias,
alternatively, are generally found in the mesopelagic where
photosynthesis cannot occur, but they can feed on sinking particles
(Stukel et al., 2018). In an imaging-based study of the whole Rhizaria
community, Biard and Ohman (2020) noted clear patterns in vertical
zonation which largely corresponded to different trophic roles. In the
oligotrophic ocean, Blanco-Bercial et al. (2022) also noted that the
protist community, including Rhizaria, partition along an autotroph and
mixotroph to heterotroph gradient with increasing depth in the water
column. Yet, few studies have made direct attempts to relate rhizaria
abundances to environmental factors (Biard and Ohman, 2020). In part,
this is due to the fact few studies have been able to sample Rhizaria in
the same location over a consistent timeframe (Boltovskoy et al., 1993;
Gutiérrez-Rodríguez et al., 2022; Hull et al., 2011; Michaels et al.,
1995; Michaels, 1988). Furthermore, no studies have utilized imaging,
arguably the best method for quantifying rhizaria, consistently
throughout the full mesopelagic. Given this lack of information, there
are many unknowns with respect to Rhizaria ecology, seasonality and
phenology across different groups.

In this study, we present a comprehensive assessment of large Rhizaria
measured for over a year from regularly occurring cruises at monthly
intervals. We utilized an in-situ imaging approach to facilitate
abundance calculations. With this dataset, we address two critical aims.
1) Quantification of large Rhizaria throughout the epipelagic (0-200m)
and mesopelagic (200-1000m) over the course of an annual cycle. These
data were collected in the Sargasso Sea, and represents the first study
of its kind in an oligotrophic system; and 2) We aim to test the
hypothesis that Rhizaria exhibit niche partitioning according to trophic
roles. This hypothesis makes several predictions, including vertical
zonation, as seen in prior studies, but also that environmental
variables related to trophic strategy will explain abundance patterns.
Specifically, autotrophic/mixotrophic taxa will correspond to variables
related to autotrophy (chl-a concentration, primary productivity, local
DO maxima) and other rhizaria will correspond to factors which promote
heterotrophy (particle concentration, flux, and local DO minima).

@@

Data were collected in collaboration with the Bermuda Atlantic
Time-series Study (Lomas et al., 2013; Michaels and Knap, 1996) on board
the R/V Atlantic Explorer. Cruises were conducted at approximately
monthly intervals. Rhizaria individuals were sampled using the
Underwater Vision Profiler 5 {[}UVP5; Picheral et al. (2010){]}, a tool
which is well established to accurately quantify large Rhizaria (Barth
and Stone, 2022; Biard et al., 2016; Biard and Ohman, 2020; Drago et
al., 2022; Llopis Monferrer et al., 2022; Panaïotis et al., 2023; Stukel
et al., 2019; Stukel et al., 2018). The UVP5 was mounted to the sampling
rosette and collected data autonomously on routine casts, from which
only the downcast data are utilized. The UVP5 was deployed from
June-September 2019 then from October 2020 - January 2022, during which
time the BATS region was sampled for 3-5 days at monthly intervals.
Casts were filtered to only include data collected in the BATS region,
far offshore of Bermuda in the Sargasso Sea (approximately
31.0\(^oN\)-32.5\(^oN\), 64.25\(^oW\)-63\(^oW\); Supplemental Figure 1).
In general, casts extended to either 200m, 500m, or 1200m deep, with a
few extended into the bathypelagic (4500m). However, Rhizaria were only
typically found in large abundances throughout the epipelagic and
mesopelagic zones. As such, we limit this study to results from the
upper 1000m of the water column.

A variety of biotic and abiotic data were collected during each BATS
cruise. Briefly, we will explain the data utilized in this study. The
UVP5 provided particle count data at a high-frequency from each cast.
Particle concentration was calculated from this data for all particles
184\(\mu m\) - 450\(\mu m\). The lower size range was set by what could
be reliably sampled by the UVP5's pixel resolution (\textgreater2px;
0.092mm per pixel) and the upper size range is representative of a
potential prey field for mesozooplankton (Whitmore and Ohman, 2021). For
each UVP cast supporting continuous profiles of the CTD parameters
salinity, temperature, and auxiliary CTD channels; Dissolved Oxygen
(DO), in-situ chlorophyll fluorescence were measured at 24Hz using the
BATS CTD package. On select casts, Niskin bottles were used to collect
bacterial abundance estimates (via epifluorescence microscopy) as well
as measure inorganic nutrients (\(NO_3\), and \(Si\) as silicate/sicilic
acid) at discrete depths. On each cruise, flux estimates of total mass,
organic carbon, and nitrogen were also collected using sediment traps;
in the present study we utilized flux to the mesopelagic as the flux at
200m. Also primary productivity was estimated through measuring
\(^{14}C\) uptake rates from in-situ incubations. For full descriptions
of the BATS sampling program and methods, see Knap et al. (1997) and
Lomas et al. (2013) for a review. Additionally, data can be viewed
online
\href{https://bats.bios.asu.edu/bats-data/}{(https://bats.bios.asu.edu/bats-data/)}.

Environmental data were processed in a variety of ways to match the
format of the Rhizaria abundance estimates (see below). CTD data were
collected at higher frequency than the UVP (24Hz vs 15Hz respectively),
so these data were averaged within matching bins to the UVP5 data. Data
from Niskin bottles were first linearly interpolated in depth at 1m
resolution then time averaged over the cruise, then subsequently
averaged into matching UVP5-sized bins. Primary productivity estimates
were totaled within the euphotic zone to represent a ``total euphotic
productivity''.

@@

Individual vignettes of Rhizaria images were identified using the
classification platform Ecotaxa (Picheral et al., n.d.). Data were
pre-sorted utilizing a random-forest classifier and pre-trained learning
set. Taxonomic classification were done based on morphology exclusively.
While there are sparse taxonomic guides for in-situ images of rhizaria,
identification largely relied on descriptions in (Nakamura and Suzuki,
2015; Suzuki and Not, 2015; and Biard and Ohman, 2020). Using the
aforementioned sources and publicly available ecotaxa projects, we
constructed a guide accessible at:
\url{https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf}.
Broadly, Rhizaria were classified as Foraminifera, Radiolarias
(Acantharea or Collodaria), or as a variety of Phaeodaria families
(Figure 1). When identification could not be confidently made between a
few candidate taxa, a less specific label was used. As a result, we have
data from ``unidentified Rhizaria'', which typically were vignettes not
distinguishable between Aulacanthidae or Acantharea or ``unidentified
Phaeodaria'', which are clearly Phaeodaria but not distinguishable into
a family.
